\vspace{2cm} \LARGE \bf Seasonal influenza circulation patterns and projections for 2017-2018

### 2017-02-17 auspice/reports/tex/feb-2017.tex
revert some of the frequency figures, fix mismatches in capitalization, add L&L

We base our primary analysis on a set of viruses collected between Jan 2015 and Feb 2017 (Fig.\ \ref{H3N2_counts}).
We analyze mutation frequency trajectories separately in different geographic regions using all available sequence data.
Our phylogenetic analysis is based on a subsample of approximately 100 viruses per month where available and seeking to equilibrate sample counts geographically where possible.
This equilibration attempts to collect equal samples from Africa, China, Europe, Japan/South Korea, North America, Oceania, South America, South Asia, Southeast Asia and West Asia.
In the following analysis we collapse samples from China, South Asia, Southeast Asia, Japan and Korea into a single region referred to here as ``Asia'', resulting in Asia possessing greater sample counts than North America or Europe.
There is reasonably broad geographic sampling throughout, except Dec 2016 and Jan 2017, which have more samples from Europe and North America.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\textwidth]{../figures/feb-2017/H3N2_tree.png}
	\caption{\textbf{H3N2 / 3c2.a phylogeny colored by genotype.}
	}
	\label{H3N2_tree}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.95\textwidth]{../figures/feb-2017/H3N2_mutations.png}
	\caption{\textbf{Frequency trajectories of 3c2.a subclades.}
	We estimate frequencies of different clades based on sample counts and collection dates.
	We use a Brownian motion process prior to smooth frequencies from month-to-month.
	Transparent bands show an estimate the 95\% confidence interval based on sample counts.
	The final point represents our frequency estimate for Feb 1 2017.
	}
	\label{H3N2_mutations}
\end{figure}

@@

The local branching index (LBI), a phylogenetic indicator of clade growth \cite{neher2014predicting}, corroborates the observations made on the basis of clade frequencies (Fig.\ \ref{H3N2_lbi}).
Over the last 6 monthes, the 131K clade has highest LBI, while the clade 171K has highest LBI when considering sequences in all of 2016.
Interpretation of the sequence based indices such as number of epitope or non-epitope mutations is difficult due the extensive genetic heterogeneity.
The clade 131K has accumulated the same number of epitope mutations as 171K but has fewer non-epitope mutations relative to A/Texas/50/2012.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\textwidth]{../figures/feb-2017/H3N2_lbi.png}
	\caption{\textbf{H3N2 phylogeny colored by local branching index.}
	}
	\label{H3N2_lbi}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\textwidth]{../figures/feb-2017/H1N1pdm_tree.png}
	\caption{\textbf{H1N1pdm / 6b.1 phylogeny colored by genotype.}
	}
	\label{H1N1pdm_tree}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.95\textwidth]{../figures/feb-2017/H1N1pdm_mutations.png}
	\caption{\textbf{Frequency trajectories of 6b.1 subclades.}
	We estimate frequencies of different clades based on sample counts and collection dates.
	We use a Brownian motion process prior to smooth frequencies from month-to-month.
	Transparent bands show an estimate the 95\% confidence interval based on sample counts.
	The final point represents our frequency estimate for Feb 1 2017.
	}
	\label{H1N1pdm_mutations}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.95\textwidth]{../figures/feb-2017/Vic_mutations.png}
	\caption{\textbf{Frequency trajectories of B/Vic variants.}
	We estimate frequencies of different clades based on sample counts and collection dates.
	We use a Brownian motion process prior to smooth frequencies from month-to-month.
	Transparent bands show an estimate the 95\% confidence interval based on sample counts.
	The final point represents our frequency estimate for Feb 1 2017.
	}
	\label{Vic_mutations}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.95\textwidth]{../figures/feb-2017/Yam_mutations.png}
	\caption{\textbf{Frequency trajectories of B/Yam variants.}
	We estimate frequencies of different clades based on sample counts and collection dates.
	We use a Brownian motion process prior to smooth frequencies from month-to-month.
	Transparent bands show an estimate the 95\% confidence interval based on sample counts.
	The final point represents our frequency estimate for Feb 1 2017.
	}
	\label{Yam_mutations}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{../figures/feb-2017/Yam_tree.png}
	\caption{\textbf{B/Yam phylogeny colored by genotype.}
	}
	\label{Yam_tree}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\textwidth]{../figures/feb-2017/Yam_lbi.png}
	\caption{\textbf{B/Vic phylogeny colored by local branching index.}
	}
	\label{Yam_lbi}
\end{figure}

>>>>

We base our primary analysis on a set of viruses collected between Jan 2015 and Feb 2017 (Fig.\ \ref{H3N2_counts}).
We analyze mutation frequency trajectories in different geographic regions using all available sequence data for each region.
Our phylogenetic analysis is based on a subsample of approximately 100 viruses per month where available and seeking to equilibrate sample counts geographically where possible.
This equilibration attempts to collect equal samples from Africa, China, Europe, Japan/South Korea, North America, Oceania, South America, South Asia, Southeast Asia and West Asia.
In the following analysis we collapse samples from China, South Asia, Southeast Asia, Japan and Korea into a single region referred to here as ``Asia'', resulting in Asia possessing greater sample counts than North America or Europe.
There is reasonably broad geographic sampling throughout, except Dec 2016 and Jan 2017, which have more samples from Europe and North America.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\textwidth]{../figures/feb-2017/h3n2_tree.png}
	\caption{\textbf{H3N2 / 3c2.a phylogeny colored by genotype.}
	}
	\label{H3N2_tree}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.95\textwidth]{../figures/feb-2017/h3n2_frequencies.pdf}
	\caption{\textbf{Frequency trajectories of 3c2.a subclades.}
	We estimate frequencies of different clades based on sample counts and collection dates.
	We use a Brownian motion process prior to smooth frequencies from month-to-month.
	Transparent bands show an estimate the 95\% confidence interval based on sample counts.
	The final point represents our frequency estimate for Feb 1 2017.
	}
	\label{H3N2_mutations}
\end{figure}

@@

The local branching index (LBI), a phylogenetic indicator of clade growth \cite{neher2014predicting}, corroborates the observations made on the basis of clade frequencies (Fig.\ \ref{H3N2_lbi}).
Over the last 6 monthes, the 131K clade has highest LBI, while the clade 171K has highest LBI when considering sequences in all of 2016.
Interpretation of the sequence based indices such as number of epitope or non-epitope mutations \cite{luksza2014predictive} is difficult due the extensive genetic heterogeneity.
The clade 131K has accumulated the same number of epitope mutations as 171K but has fewer non-epitope mutations relative to A/Texas/50/2012.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\textwidth]{../figures/feb-2017/h3n2_lbi.png}
	\caption{\textbf{H3N2 phylogeny colored by local branching index.}
	}
	\label{H3N2_lbi}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\textwidth]{../figures/feb-2017/h1n1pdm_tree.png}
	\caption{\textbf{H1N1pdm / 6b.1 phylogeny colored by genotype.}
	}
	\label{H1N1pdm_tree}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.95\textwidth]{../figures/feb-2017/h1n1pdm_frequencies.pdf}
	\caption{\textbf{Frequency trajectories of 6b.1 subclades.}
	We estimate frequencies of different clades based on sample counts and collection dates.
	We use a Brownian motion process prior to smooth frequencies from month-to-month.
	Transparent bands show an estimate the 95\% confidence interval based on sample counts.
	The final point represents our frequency estimate for Feb 1 2017.
	}
	\label{H1N1pdm_mutations}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.95\textwidth]{../figures/feb-2017/vic_frequencies.pdf}
	\caption{\textbf{Frequency trajectories of B/Vic variants.}
	We estimate frequencies of different clades based on sample counts and collection dates.
	We use a Brownian motion process prior to smooth frequencies from month-to-month.
	Transparent bands show an estimate the 95\% confidence interval based on sample counts.
	The final point represents our frequency estimate for Feb 1 2017.
	}
	\label{Vic_mutations}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.95\textwidth]{../figures/feb-2017/yam_frequencies.pdf}
	\caption{\textbf{Frequency trajectories of B/Yam variants.}
	We estimate frequencies of different clades based on sample counts and collection dates.
	We use a Brownian motion process prior to smooth frequencies from month-to-month.
	Transparent bands show an estimate the 95\% confidence interval based on sample counts.
	The final point represents our frequency estimate for Feb 1 2017.
	}
	\label{Yam_mutations}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{../figures/feb-2017/yam_tree.png}
	\caption{\textbf{B/Yam phylogeny colored by genotype.}
	}
	\label{Yam_tree}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\textwidth]{../figures/feb-2017/yam_lbi.png}
	\caption{\textbf{B/Vic phylogeny colored by local branching index.}
	}
	\label{Yam_lbi}
\end{figure}
